---
title: "Project 2 - Data 607 - Michael Hayes"
author: "Michael Hayes"
date: "March 8, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(dplyr)
```

## Color and Heat Absorption Data

For this "tidying" we will utilize Christopher Ayre's example dataset of Heating and Cooling Absorption.  

Discussion board can be found here: https://bbhosted.cuny.edu/webapps/discussionboard/do/message?action=list_messages&course_id=_1705328_1&nav=discussion_board&conf_id=_1845527_1&forum_id=_1908779_1&message_id=_31283025_1

### Step 1 - Load Dataset

Christopher provided the data in a .csv file, which I've uploaded to github:

```{r}
heating <- read.csv("https://raw.githubusercontent.com/murphystout/data-607/master/heating_cooling.csv")

head(heating)
```

Christopher adeptly pointed out several issues with the data set that make it "untidy".  These are:

1. The variable for "time elapsed" does not have its own column.  In this case we see a column for each ten minute interval.  These should be collapsed into one column.

2. Each color should have its own column.  We will treat one "observation" as the temperature across all the colors, and so the color is merely a variable for a single observation, hence they should all be placed on a single row.

3. Multiple observational units are observed in the same table.  In particular the "heating" and "cooling" data are in one table.

After we have tidied it up, we will do some exploratory analysis and visualizations on the data.

### Step 2 - Gather Times

For this step we will gather multiple timestamp columns into a single column.

As a intermediate step, let's first rename the columns to that they take on a numerical value, this will help with plotting the value later on.

```{r}
colnames(heating) <- c("color",0,10,20,30,40,50,60,"phase")

heating <- gather(heating, time, temperature, "0":"60")

head(heating, 20)

```

Finally, let's make sure the time is stores as numeric values:

```{r}
heating$time <- as.numeric(heating$time)

```


### Step 3 - Spreading Colors to Columns

Now that's we've gathered up the temperature columns, let's spread out the color columns to include one temp reading for each color + timestamp combination.


```{r}
heating <- spread(heating, color, temperature)

heating

```

### Step 4 - Split observational units to separate tables.

Thankfully this is as simple as subsetting the data based on the "phase" column.  I saved this for the last step to save us from having to perform the tidying operations twice.

```{r}
cooling <- subset(heating, phase == "cooling")
heating <- subset(heating, phase == "heating")

head(cooling)
head(heating)

```

### Step 5 - Exploraty Data Analysis

Let's plot these data in a line graph to get a visual representation of how the colors responded to heating and cooling:

```{r}

plot(x = heating$time, y = heating$black, type = "l", col = "black", xlab = "Time Elapsed (Minutes)", ylab = "Temp (Farheneit)", main = "Heating/Color Absorption")
lines(x = heating$time, y = heating$red, col = "red")
lines(x = heating$time, y = heating$green, col = "green")
lines(x = heating$time, y = heating$pink, col = "pink")
lines(x = heating$time, y = heating$white, col = "grey")

plot(x = cooling$time, y = cooling$black, type = "l", col = "black", xlab = "Time Elapsed (Minutes)", ylab = "Temp (Farheneit)", main = "Cooling/Color Asborption")
lines(x = cooling$time, y = cooling$red, col = "red")
lines(x = cooling$time, y = cooling$green, col = "green")
lines(x = cooling$time, y = cooling$pink, col = "pink")
lines(x = cooling$time, y = cooling$white, col = "grey")

```

### Initial findings and Next steps

The graphs look neat, and we can see that black is the fastest heat absorber.  

The graphs also look symmetrical, but now that we see it in this form, it might make sense to view the cooling and heating data in one graph.  

However, the data requires a bit more finagling to get this correct, such as:

1: Minute "60" of the Heating Data is equivalent of Minute "0" of the Cooling.

2: Minutes elapsed in the Cooling data need to be increased by 60 in order to create one continuous time series.

Let's do it!

### Combining Heating and Cooling Data

```{r}
# Remove minute 0 of the cooling dataset:

heat_cool <- cooling[-1,]

# Add 60 to the time column.

heat_cool$time <- as.numeric(heat_cool$time) + 60

# Stack this underneath the heating data

heat_cool <- rbind(heating, heat_cool)

heat_cool

```

Now we have a nice, neat and tidy dataset showing heating and cooling times.  Let's revisit those graphs we generated previously:


```{r}
plot(x = heat_cool$time, y = heat_cool$black, type = "l", col = "black", xlab = "Time Elapsed (Minutes)", ylab = "Temp (Farheneit)", main = "Cooling/Color Asborption")
lines(x = heat_cool$time, y = heat_cool$red, col = "red")
lines(x = heat_cool$time, y = heat_cool$green, col = "green")
lines(x = heat_cool$time, y = heat_cool$pink, col = "pink")
lines(x = heat_cool$time, y = heat_cool$white, col = "grey")

```

#### Calculating Rates (Hourly)

Let's get a bit more quantitative.  Let's calculate the rates of heating and cooling for each of the colors:

```{r}
heating_rate <- (heating[7,3:7] - heating[1,3:7])/60

heating_rate


cooling_rate <- (cooling[7,3:7] - cooling[1,3:7])/60

cooling_rate
```

Since the starting and ending temperatures were equivalent, we see the overall heating and cooling rates to be symmetrical to one another.  

According to this test, a colors heating rate also dicates its cooling rate (or heat retention), at least on average over 120 minutes.

Looking at these visually:

```{r}
heating_rate <- gather(heating_rate, color, temp)

barplot(heating_rate$temp, col = heating_rate$color, names.arg = heating_rate$color, main = 'Heating Rates (by Color)', xlab = "Color", ylab = "Rate (Degrees per minute)")

```

However, this is looking at averages over the hour.  But what does temp change look like within each 10 minute interval?

We can find this programmatically:

### Calculating Rates (Per 10 Minutes)

```{r}
black_ht <- diff(heating$black)/10
green_ht <- diff(heating$green)/10
pink_ht <- diff(heating$pink)/10
red_ht <- diff(heating$red)/10
white_ht <- diff(heating$white)/10

black_cl <- diff(cooling$black)/10
green_cl <- diff(cooling$green)/10
pink_cl <- diff(cooling$pink)/10
red_cl <- diff(cooling$red)/10
white_cl <- diff(cooling$white)/10

ht_rates <- data.frame(black_ht, black_cl, green_ht, green_cl, pink_ht, pink_cl, red_ht, red_cl, white_ht, white_cl)

ht_rates

```


Let's take a look at these visually:

```{r}
plot(x = seq(10, 60, 10), y = black_ht, type = "l", col = "black", ylim = c(-1.5,1.5), main = "Heating and Cooling Rates", sub = "Postive Values are Heating Rates, Negative are Cooling Rates", xlab = "Time Elapsed (Minutes)", ylab = "Heating and Cooling Rates")
lines(seq(10, 60, 10),y = black_cl, col = "black")
lines(seq(10, 60, 10),y = green_ht, col = "green")
lines(seq(10, 60, 10),y = green_cl, col = "green")
lines(seq(10, 60, 10),y = pink_ht, col = "pink")
lines(seq(10, 60, 10),y = pink_cl, col = "pink")
lines(seq(10, 60, 10),y = red_ht, col = "red")
lines(seq(10, 60, 10),y = red_cl, col = "red")
lines(seq(10, 60, 10),y = white_ht, col = "grey")
lines(seq(10, 60, 10),y = white_cl, col = "grey")
```

This chart shows both heating and cooling rates.  The heating rates are postive (top of chart), while the cooling rates are negative (bottom of chart). 

Matching like colors can show you how that color behaved in its heating and cooling phase.

Being that our ultimate averages were very symmetrical (i.e. over the full 120 minute span), we might expect that each smaller interval would be symmetrical too.  

However that doesn't always to be the case in this data.  Note the green line is often twice the magnitude of its counterpart.

We can also see that all colors tend to converge to low values at the end of *both* periods.  Perhaps this speaks to a type of heating saturation paired with a similar flatline of cooling.

### Some Conclusions and Questions for Future Analysis

Some initial conclusions from our exploratory data analysis:

1. Black has the fastest heating rate, and ~0.72 degrees per minute.  White has the slowest heating rate, at ~0.33 degrees per minute.  This was probably suspected based on known heuristics, and the data seems to confirm it.

2. Heating rates and cooling rates were symmetrical over a 120 minute span.  However, they don't seem to be symmetric over smaller 10 minute spans.  Lots of variation of rates across that time.

Some questions it raised:

1. There seems to be a wide varience for temperature changes in the 10 minute intervals.  Is this typical?  Do temperature changes "slow" or otherwise change during based on when they occur in the time series?

2. Heat absorbtion may very well not be a linear activity, a perhaps more detailed detail in needed to really understand the dynamics of these rates.



