---
title: "Week 3 Assignment - Data 607"
author: "Michael Hayes"
date: "February 14, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(stringr)

raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"

name <-unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name

phone <- unlist(str_extract_all(raw.data, "\\(?(\\d{3})?\\)?(-| )?\\d{3}(-| )?\\d{4}"))
phone

```


Copy the introductory example. 

```{r}

name
```

a. The vector name stores the extracted names. Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name. 

```{r}
#Determine which names *don't* follow first name + last name (i.e. user lastname,firstname)

last_first <- name[str_detect(name,"[[:alpha:]],")]

last_first

#Split members on the comma, swap via backreferencing

first_last <- str_replace(last_first,"([^,]+),\\s([^,]+)","\\2 \\1")

first_last

#Place back into original vector.

name[str_detect(name,"[[:alpha:]],")] <- first_last
```

b. Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.). 

```{r}
title <- str_detect(name, "[[:alpha:]]{2,3}[.]")

title

name[title]



```

c. Construct a logical vector indicating whether a character has a second name.

```{r}
#Assuming second name means middle name OR last name

#Remove titles, then check for a space.  This would indicate there are 2 words, hence 2 names.

second_name <- str_detect(str_replace(name,"[[:alpha:].]{2,3}\\. ",""),"\\s")

second_name 

name[second_name]

#If they are referring to a middle name, do the same, just checking for 2 spaces, hence 3 words:

middle_initial <- str_detect(str_replace(name,"[[:alpha:].]{2,3}\\. ",""),"\\s.*\\s")

middle_initial

name[middle_initial]


```


Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression. 

NOTE: I've included examples that return false in order to better exemplify the limits of the pattern.

[0-9]+\\$ 

1 or more digits followed by a dollar sign.

```{r}
answer_4a <- c("$1000", "1000", "100$", "1000$")

str_detect(answer_4a, "[0-9]+\\$")


```


\\b[a-z]{1,4}\\b 

Word break follow by 1 to 4 lowercase letters followed by another word break.

```{r}
answer_4b <- c("This is a full sentence with several one to four letter words")

str_extract_all(answer_4b,"\\b[a-z]{1,4}\\b")


```



.*?\\.txt$ 

Zero or more of any chacter (optionally), ending with a literal period and 'txt' 


```{r}
answer_4c <- c("filename.txt", ".txt", "This sentence doesn't end with a .txt extension")

str_detect(answer_4c, ".*?\\.txt$")

```

\\d{2}/\\d{2}/\\d{4} 

2 digits followed by a foreward slash followed by 2 digits followed by a forward slash followed by 4 digits 

A common date notation.

```{r}

answer_4d <- c("04/04/2019", "12/23/2011", "99/99/9999", "1/2/3432")

str_detect(answer_4d, "\\d{2}/\\d{2}/\\d{4}")

```


<(.+?)>.+?</\\1>

Less than sign folowed by an optional 1 or more chacters followed by a greater than sign, followed by an optional 1 or more chacters, followed by a less than sign, a foreward slash, and the first matching group, and a final greater than sign.

I.e.: This will extract HTML tags and their containments.

```{r}


str_extract_all("This is HTML markup code, with <strong>bolded text</strong> as well as <em>italics</em>", "<(.+?)>.+?</\\1>")


```


9.The following code hides a secret message. Crack it with R and regular expressions. Hint: Some of the characters are more revealing than others! The code snippet is also available in the materials at www.r-datacollection.com.

```{r}

secretcode <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"

str_replace_all(secretcode,"[[:lower:]]|[[:digit:]]","")

```

